Projecting Global Filesystems into Local Containers for High-throughput Computing on HPC Resources

### 2019-01-23 paper.tex
Updates from ShareLaTeX
Added curated
Update to paper state

Shrinkwrap provides a mechinsm for projecting parts of CVMFS,
but is not by itself a complete solution as it does not integrate with job submission or manage images after creation.
In the case of a small number of manually created container images,
this is not a serious problem.
Administrators could manage a small collection of images in this case and adjust jobs to work with their setup.
Thus for integrating with the existing systems at NERSC,
Shrinkwrap alone might be sufficient.
To support CVMFS as a site-wide service with a potentially varied user base and workload,
a more automated solution
(e.g.\ a batch system plugin that can automatically prepare container images on job submission)
is required.
This both minimizes direct administrator burden and more closely approximates user experience on other resources like the WLCG.
Our first attempt at such a system,
described in the next sections,
took the simplest possible approach to image.
As we were developing this approach, however,
a number of issues in real-world usage led us to conculde that a somewhat more sophisticated solution was required.

@@

The simplest approach for automatically projecting container images is to create a suitable image on each batch job submission.
Unfortunately, this solution is very wasteful of computing time and storage space.
Even a minimal image consisting of a precise set of required libraries and frameworks ranges in the tens of gigabytes,
depending on the application.
The full-repo images being tested at NERSC are multiple terabytes in size and take the majority of a day to create.
In addition, image creation must be completed before any actual processing occurs,
i.e.\ there is no way to start with a partial image that supports the first steps of a pipeline and finish building the image as the pipeline runs.
There is also the cost of transferring and storing freshly created images on any nodes that will use them,
further delaying job dispatch and increasing load on the network and storage infrastructure.

@@

After creating a set of static images,
the next step is to transfer them to local caches on the workers.
Worker nodes are assumed to have some fixed amount of storage available for caching images.
If the size of the ``working set'' of images is smaller than the node-local storage,
then each worker can keep a complete set of images.
While ideal, this case is not necessarily realistic beyond a few images.
With small/medium images in the tens-to-hundreds of gigabytes range,
maintining a complete set of projected CVMFS images at each worker requires substantial resource provisioning.
Thus it is important for workers be able to to cache only a subset of the available projected images.
There are numerous caching policies and options available,
but choice of worker cache strategy is not critically important for this application.
Since projections and images of CVMFS can be re-created on demand,
cache evictions might result in computation overhead but would not otherwise affect the functionality or correctness of the system.
In the absence of a compelling and site-specific reason,
a simple caching policy such as LRU would be sufficient.
As we will see in section ??,
LRU will also be an important piece of our final proposed approach.

Under this design,
on job submission the bridge machine first checks if the same projection had been requested before.
If so, it arranges to transfer its cached container image to worker nodes that lack that image locally.
If the bridge node does not have that projection built already,
it uses Shrinkwrap to create the requested projection,
then invokes mksquashfs to create the container image.
Once the bridge node places the newly created image in its cache,
the process proceeds as in the previous case.
The architecture of this initial approach is shown in figure ???.

%Node-local scratch storage is sufficient for temporarily caching these images.
%We also assumed that durable storage is available on the bridge machine for storing the data cache and projections associated with Shrinkwrap,
%as well as a cache of container images derived from these projections.
%Again, we chose LRU as a caching policy for our evaluation,
%though administrators would be free to experiment with different policies.

%Technologies like Docker currently operate this way (often to the chagrin of users and administrators doing scientific computing).

@@

@@

The first step in simulating system behavior over time was generating synthetic requests for images.
We tested two schemes for generating image requests taking into account the properties of the software collections,
as well as recorded user activity patterns.
We also generated images via a uniform random scheme for comparison.
We consulted with the developers of CVMFS,
as well as HEP researchers at the University of Notre Dame collaborating with CERN to determine how current users interact with CVMFS.
Since CVMFS is not currently widely deployed on HPC resources,
the prospective user base for our container management system is limited.
We were, however, informed of some researchers and groups experimenting with static mirrors of CVMFS for container images,
including projects at NERSC and Stanford University,
with one of the setups available publicly at \url{https://github.com/wyang007/atlas-fat-container}.

@@

@@

\begin{figure*}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_cache_percent_plt.pdf}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_cache_percent_plt.pdf}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_request_percent_plt.pdf}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_request_percent_plt.pdf}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_uses_percent_plt.pdf}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_uses_percent_plt.pdf}
\caption{}
\end{figure*}

>>>>

Shrinkwrap provides a mechanism for projecting parts of CVMFS,
but is not by itself a complete solution as it does not integrate with job submission or manage images after creation.
In the case of a small number of manually created container images,
this is not a serious problem.
Administrators could manage a small collection of images in this case and adjust jobs to work with their setup.
Thus for integrating with the existing systems at NERSC,
Shrinkwrap alone might be sufficient.
To support CVMFS as a site-wide service with a potentially varied user base and workload,
a more automated solution
(e.g.\ a batch system plugin that can automatically prepare container images on job submission)
is required.
This both minimizes direct administrator burden and more closely approximates user experience on other resources like the WLCG.
Our first attempt at such a system,
described in the next sections,
took the simplest possible approach to image.
As we were developing this approach, however,
a number of issues in real-world usage led us to conculde that a somewhat more sophisticated solution was required.

\begin{figure}
\includegraphics[width=\linewidth]{curated/cdf/cdf.pdf}
\label{fig:cdf}
\if 0
lots of small files
table shows duplication
long tail of large files
different scales, similar usage across repos

\fi
\end{figure}

\begin{figure}
\includegraphics[width=\linewidth]{curated/closure/closure.pdf}
\label{fig:picks}
\if 0
you get more than you ask for
smaller requirements have relatively large responses
non-linear, so bigger ones have more duplicates
appraoches repo size in limit
merging is subject to drop-off(more duplication)
\fi
\end{figure}

To evaluate the effect of including dependencies in container images,
we made a large number of random selections of packages from the SFT repo and then compared with the set of dependencies required by each selection.
Figure~\ref{fig:picks} shows the results over a large number of selection sizes.
We observe of course that selection size increases uniformly along the $x$ axis.
The on-disk size of the selections appears to increase proportionally to selection size.
Recursively including the dependencies of these same selections,
however, results in a significant increase in number of packages and size on disk.
For small selections (less than 100~packages),
the complete images with dependencies might contain 5x as many packages as requested,
with storage increasing proportionally.
With larger selections, however,
this increase becomes less prominent.
This non-linear change is the result of the tree structure of the software dependencies.
There are a number of core components that are transitive dependencies of a large number of packages.
These common dependencies are only included once,
so subsequent selections that depend on them do not add them again.
As the number of selections increases,
this curve will approach the total repo size.
Also note that the non-linear increase in size with increasing selection size means that the size of combining two selections of 100 packages (and their dependencies) is not as simple as adding them together.
Due to duplication, we would expect the resulting combined size to be lower,
as if choosing 200 packages from the start.

\if 0
spread across everyone, lot of commonality/overlap
few with very high usage, very large number of rarely used
lot not observed
\fi

Since there are currently few regular CVMFS on HPC users,
we decided to simulate the activity of users requesting container images.
We decided to simulate at the granularity of packages rather than individual files,
because this results in more robust container images.
As mentioned previously,
fine-grained tracing runs the risk of leaving out important dependencies.
In addition, working only based on the properties of individual files ignores the structure inherent in software collections.
Looking only at the properties of individual files,
Figure~\ref{fig:cdf} shows cumulative distributions for file sizes in several of the largest and most active CVMFS repos.
The repos vary in scale,
but all show similar usage patterns.
We see that across all repos,
the majority of files are small (between 1~KB and 1~MB).
The repos do contain, however,
a small number of large files.
This fits our expectation that the CVMFS repos primarily contain software.
Most individual files are small,
but the repo size is defined by the large number of packages in the software collection.

We tested two schemes for generating image requests taking into account the properties of the software collections,
as well as recorded user activity patterns.
We also generated images via a uniform random scheme for comparison.
We consulted with the developers of CVMFS,
as well as HEP researchers at the University of Notre Dame collaborating with CERN to determine how current users interact with CVMFS.
Since CVMFS is not currently widely deployed on HPC resources,
the prospective user base for our container management system is limited.
We were, however, informed of some researchers and groups experimenting with static mirrors of CVMFS for container images,
including projects at NERSC and Stanford University,
with one of the setups available publicly at \url{https://github.com/wyang007/atlas-fat-container}.

@@

The simplest approach for automatically projecting container images is to create a suitable image on each batch job submission.
Under this design,
on job submission the bridge machine first checks if the same projection had been requested before.
If so, it arranges to transfer its cached container image to worker nodes that require it.
If the bridge node does not have that projection built already,
it uses Shrinkwrap to create the requested projection,
then invokes mksquashfs to create the container image.
Once the bridge node places the newly created image in its cache,
the process proceeds as in the previous case.
The architecture of this initial approach is shown in figure ???.

Unfortunately, this solution is very wasteful of computing time and storage space.
Even a minimal image consisting of a precise set of required libraries and frameworks ranges in the tens of gigabytes,
depending on the application.
The full-repo images being tested at NERSC are multiple terabytes in size and take the majority of a day to create.
In addition, image creation must be completed before any actual processing occurs,
i.e.\ there is no way to start with a partial image that supports the first steps of a pipeline and finish building the image as the pipeline runs.
There is also the cost of transferring and storing freshly created images on any nodes that will use them,
further delaying job dispatch and increasing load on the network and storage infrastructure.

@@

@@

\begin{figure}
\includegraphics[width=\linewidth]{curated/time-series/time-series.pdf}
\label{fig:series}
\if 0
zoomed out?
cache limit, deletes start
merge tracks writes, it's the dominant factor
hits grow gradually as cache merges and fills
deletes help prune useless stuff, remaining should be frequent hits
far ends are either insert or merge, middle alphas have both
\fi
\end{figure}

Figure~\ref{fig:series} shows a single simulation of the image management system,
with several notable details.
First, we see that most of the operations are merges.
This makes sense, as this simulation was at fairly high $\alpha$.
The total bytes written also closely tracks merges,
indicating that merging is the dominant source of I/O.
We can still see inserts over the course of the simulation.
At more extreme $\alpha$ values,
we would expect to see one of these operations almost exclusively.
As the data in cache continues to rise,
we eventually see the cache limit,
after which the delete count starts increasing.
Over the course of the simulation,
inserts and deletes are filling and emptying the cache such that it remains close to its storage limit.
We also observe the number of cache hits continue to rise despite deletions.
As we will see, merging allows for a greater proportion of hits even if the amount of data remains constant.
This is because frequently used data can be merged,
reducing duplication.
The cache limit then ensures that infrequently used parts are eventually removed.

% explain alpha plot

Sweeping over the range of $\alpha$ values in this way,
we can immediately see differences in the frequency of simulated operations.
Figure~\ref{fig:opcounts} shows the upper range of $\alpha$ values where behavior differences appear.
From the lower values on the left,
the insert and delete counts are the primary (or only) operations,
with number of hits relatively constant.
This corresponds to a simple LRU-based cache.
The insert count is higher due to cache filling,
but the two tend to increase in lockstep.
As $\alpha$ increases,
image merges become more probable.
The merge count in the figure steadily increases throughout most of the upper range,
while inserts and deletes decrease.
This suggests that at high $\alpha$ values,
the cache space would be more efficiently used,
with some of the duplication merged out.
In the extreme case with $\alpha=1$,
every request is merged into a single image in cache,
hence the sudden increase in hit rate and decrease in total merges at the far right of Figure~\ref{fig:opcount}.
Note that as merging increases,
the proportion of hits generally also increases.
This indicates that more of the requests can be satisfied by previously merged images in the cache.
From the perspective of the system,
this is beneficial as there is no extra time or compute cost for handling a larger proportion of images.

\begin{figure}
\includegraphics[width=\linewidth]{curated/comparative/operation_count.pdf}
\label{fig:opcounts}
\if 0
inserts and merges switch as alpha goes up
more hits with merging
allows better responsiveness? image is immediately ready
inserts and deletes go to zero, so storage use will taper off
drop off: we keep everything, so 500 requests get merged; hit count goes up correspondingly since everything is in cache
not gonna be close to cache limit
inserts and deletes track: corresponds to LRU, inserts while filling then balanced
\fi
\end{figure}

@@

\begin{figure}
\includegraphics[width=\linewidth]{curated/comparative/distribution_efficiency.pdf}
\label{fig:dist-eff}
\if 0
explain what happened
metrics: why we care about efficiency?
cache efficiency: gauge duplication (waste), keeping working set costs lots of storage
container efficiency: unexpected overhead for each request; compute, network, etc.
place limits based on site, conditions
limits exist, we won't tell you
maybe just buy more cache?
large scale (10-100 TB), no can do...so need to merge sometimes
why 100 and 20 on the left
why they switch places: extreme cases
strike a balance
one curve might be more important for you
effects are not uniform/balanced gentle drop, massive spike in overhead on right
\fi
\end{figure}

When sweeping over the range of $\alpha$ values,
there are a large number of metrics available to summarize each simulation run.
Many, however, are highly coupled with the particular system configuration and difficult to compare as we vary the paramaters of the simulations.

We therefore chose to define two metrics, cache efficiency and container efficiency,
to indicate the relative utilization of the system.
Figure~\ref{fig:dist-eff} plots these two metrics for a representative simulation.
We defined cache efficiency as the ratio of unique data to total data in the cache.
In our case, this is equivalent to the ratio of the size of the unique packages to the total cache size.
If many images contain copies of the same packages,
the cache efficiency decreases.
This metric captures the extent of duplication within the cache across all images.
With no merging there is a high degree of duplication,
so the cache efficiency is low.
On the other end of the spectrum,
maintaining a single, large image containing all data results in cache efficiency of 100\%,
as the entire cache resides in that single image with no duplication

We defined container efficiency as the ratio of the size of the requested container
(in our case a set of requested packages plus all dependencies)
to the size of the container the system actually used for the job.
In the absence of merging,
these two are equal so the container efficiency is 100\%;
jobs are run with exactly what was requested.
By merging to allow for image reuse,
we include additional, unrequested data in container images.
The container efficiency measures this difference between requests and containers.
In the extreme case of $\alpha=1$ with a single large image, for example,
the container efficiency is poor because the entire repo is used for every request,
regardless of size.

These two extreme cases, no merging among many images and a single merged image,
can both be useful in some situations.
If the requirements of all jobs are known in advance and worker storage and bandwidth are sufficient,
maintaining a single, large image for all jobs on all workers is an useful strategy.
After the initial (costly) image creation and transfer,
there are no continuing costs or delays.
When resources are limited or requirements change regularly,
however, this approach becomes prohibitively expensive.
The high compute and bandwidth overhead apply for every image update,
which in the worst case could be every job.

The other extreme situation,
simply caching requests with no merging,
can also be viable.
This approach is simpler and does not come with extreme per-job overhead.
In addition, frequent changes to the jobs and requirements do not degrade the efficiency of each job.
At large scale, however,
the overall system efficiency suffers.
Due to duplication among images,
larger caches contain less and less unique data.
Thus to support a given repo,
it can become necessary to provision a cache much larger than the size of the repo.
With common CVMFS repos consuming several terabytes of storage each,
the amount of cache space used can quickly grow.

These two limiting factors,
the compute and transfer cost in the highly merged case and the cache efficiency in the unmerged case,
serve as limits on the viable range of $\alpha$ values for the system.
Figure~\ref{fig:dist-eff} shows two vertical lines indicating these limits on a plot of efficiencies.
The left line shows a lower limit of around 30\% on the cache efficiency.
Choosing $\alpha$ too low results in cache efficiency reaching below 20\%.
On the right, the line gives an upper limit on the likelihood of merging.
As shown in Figure~\ref{fig:opcounts},
the amount of I/O and compute to update images becomes much larger if $\alpha$ is set too high.
These two limits define a range of viable values for $\alpha$.
There is no general rule for the placement of these limits,
wh depends strongly on the performance characteristics of the execution environment,
as well as the priorities of the administrators in optimizing the system.

\begin{figure}
\includegraphics[width=\linewidth]{curated/comparative/cache_efficiency.pdf}
justification for cache efficiency
unique vs total
full x range?
merging can give more unique data in less cache space
\end{figure}

\begin{figure}
\includegraphics[width=\linewidth]{curated/comparative/container_efficiency.pdf}
\if 0
related to container efficiency: not exact but shows it
too much jitter from individual requests, so cumulative
crossover: hit rate allows reuse, so slightly less bytes written than requested
requested is based on random stream, constant
actual written goes up fast with alpha, so more overhead
proxy for compute/network/etc.
really steep at high alpha
\fi
\end{figure}

\begin{figure}
\includegraphics[width=\linewidth]{curated/comparative/distribution_efficiency_comp.pdf}
\if 0
different data patterns cause different effects
random has high utilization
pushed right, limited opportunity for merge, since no overlap
merging doesn't help
random is not realistic though
merging is not general solution, but for trees of software, look at this

tree has more duplication, starts merging much sooner
closure is very non-controvertial, but results in serious changes in efficiency
common to any software, so opportunity for optimization
common core that can merge to avoid dups
including usage stats (dist) amplifies these somewhat
real usage should be stronger than dist

use dist from now on because reasonable approximation to actual usage
similar to tree anyway
mild but noticable difference from tree, more natural approximation to real usage
\fi
\end{figure}

@@

\begin{figure*}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_cache_percent_plt.pdf}
\hfill
\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_cache_percent_plt.pdf}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_request_percent_plt.pdf}
\hfill
\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_request_percent_plt.pdf}
\includegraphics[width=0.48\linewidth]{curated/sensitivity/container_efficiency_uses_percent_plt.pdf}
\hfill
\includegraphics[width=0.48\linewidth]{curated/sensitivity/cache_efficiency_uses_percent_plt.pdf}
\if 0
[t] "just buy more cache"
efficiency is worse, since cache holds on to lots of questionable data
loads of opportunities to try merges, so we see a lot of unnecessary merges
less force removing bad choices, so more cruft can accumulate in many images
better cache utilization with tiny cache, but of course more thrashing
at scale, bad idea
large cache is too pricey at scale
[m] varying reuse doesn't affect too much
below 100 doesn't seem full/steady state yet
cache isn't settled? so haven't dropped bad merges; more unique data in cache boosts efficiency
[b] high number of requests, repeats don't matter
unique data dominates
hard to tell difference between [m] and [b], so not too sensitive to these
once you hit steady state, jobs count/repeats has little effect on efficiency
general trends of merging, not interesting by themselves
\fi
\end{figure*}

@@
